Racets: Embedding Policy{-}Agnostic Programming in Racket

### 2018-07-22 paper/paper.tex
adding some text to sec 4

%% To color math, put it in an \mc command, e.g., \mc{$a$}
\newcommand{\colorMATH}{black!20!blue}
\newcommand{\colorCOMM}{0.133,0.545,0.133}
\newcommand{\colorFAC}{black!20!red}
\newcommand{\blue}[1] {{\color{\colorMATH} #1}}
\newcommand{\fcol}[1] {{\color{\colorFAC} #1}}
\newcommand{\bmth}[1] {{\color{\colorMATH} $#1$}}
\newcommand{\code}[1]{\lstinline{#1}}

@@

\zw{Facet construction allows the users to construct a facet value that
consists of a policy, a private/positive variable that can be a basic value
or another facet value but is hidden away from other users, and a public/negative 
variable. }

\zw{Label creation dynamically creates a label binding to a given policy.
A policy consists of a policy variable and a policy body/expression. 
When a lable is invoked in the faceted executions, the policy will be called
to evaluate the given faceted value. When the policy is invoked, the policy
body (e.g. lambda function) evaluates to a boolean, which indicates whether
to observe the positive or the negative branch of a faceted value. This allows
the users to protect their data via choosing what data to be seen by others and 
hiding the sensitive data.}

\zw{Facet observation }

@@

      \end{array}
    \end{displaymath}
    \\
    \hfill \textit{(Label Creation / Observation)}
    \\
    \[
    \begin{array}{cc}
      {\color{\colorFAC}
        \inferrule[\textsc{Let-Label}]
          {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \langle \lambda x. ~ e, \rho' \rangle \\\\
           \alpha \not\in dom(\sigma') \and \sigma'' = \sigma' [\alpha \mapsto \langle \lambda x. ~ e, \rho' \rangle] \\\\
           e_2, \rho [ \ell \mapsto \alpha ], \sigma'' \Downarrow_{pc}^E \sigma''', v}
          {(\code{let-label} ~ \ell ~ e_1 ~ e_2), \rho, \sigma \Downarrow_{pc}^E \sigma''', v}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{Obs}]
                    {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \ell
                      \and e_2, \rho, \sigma' \Downarrow_{pc}^E \sigma'', v  \\\\
                      \langle (\lambda x.~e), \rho' \rangle = \sigma''(\ell) \\\\
                      e, \rho' [x \mapsto v], \sigma'' \Downarrow_{pc}^E \sigma''', v^{\pm} \\\\
                      e_3, \rho, \sigma''' \Downarrow_{pc}^E \sigma'''', v' \\\\
                      v'' = \fcol{\var{obs}(\ell, v', v^{\pm})}}
                    {(\code{obs} ~ e_1 ~ e_2 ~ e_3), \rho, \sigma \Downarrow_{pc}^E \sigma'''', v''}}
    \end{array}
    \]
    \\
    \hfill \textit{(Possibly-Faceted Application)} \boxed{\fcol{(v_1 ~ v_2), \rho, \sigma \Downarrow_{pc}^A \sigma, v}}
    \\
    \begingroup
    \color{\colorMATH}
    \begin{displaymath}
      \begin{array}{c}
        \begin{array}{ccc}
          {\color{\colorMATH}
          \inferrule[\textsc{App-$\star$}]
          { }
          {(\star ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma, \star}}
          & 
          {\color{\colorMATH}
          \inferrule[\textsc{App-Base}]
          {e, \rho'[x \mapsto v], \sigma \Downarrow_{pc}^E \sigma', v' }
          {(\langle \lambda x.~e, \rho' \rangle ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{App-Split}]
          {\{+\ell, -\ell\} \cap pc = \varnothing \and
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc \cup \{+\ell\}}^A \sigma', {v^+}' \\\\
            (v^- ~ v), \rho, \sigma' \Downarrow_{pc \cup \{-\ell\}}^A \sigma'', {v^-}' \and
            v' = \formfacet{\{+ \ell\}}{{v^+}'}{{v^-}'}}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma'', v'}}
        \end{array}
        \\
        \begin{array}{cc}
        \\\\
          {\color{\colorFAC}
          \inferrule[\textsc{App-Facet-Pos}]
          {+\ell \in pc \\\\
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{App-Facet-Neg}]
          {-\ell \in pc \\\\
            (v^- ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
        \end{array}
      \end{array}
    \end{displaymath}
    \endgroup
    \label{fig:semantics}
\end{figure*}

@@

>>>>

%% To color math, put it in an \mc command, e.g., \mc{$a$}
\newcommand{\colorMATH}{black!20!blue}
\newcommand{\colorCOMM}{0.133,0.545,0.133}
\newcommand{\colorFAC}{black!20!red}
\newcommand{\blue}[1] {{\color{\colorMATH} #1}}
\newcommand{\fcol}[1] {{\color{\colorFAC} #1}}
\newcommand{\bmth}[1] {{\color{\colorMATH} $#1$}}
\newcommand{\rmth}[1] {{\color{\colorFAC} $#1$}}
\newcommand{\code}[1]{\lstinline{#1}}

@@

@@

      \end{array}
    \end{displaymath}
    \\
    \hfill \textit{(Label Creation / Observation)}
    \\
    \[
    \begin{array}{cc}
      {\color{\colorFAC}
        \inferrule[\textsc{Let-Label}]
          {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \langle \lambda x. ~ e, \rho' \rangle \\\\
           \alpha \not\in dom(\sigma') \and \sigma'' = \sigma' [\alpha \mapsto \langle \lambda x. ~ e, \rho' \rangle] \\\\
           e_2, \rho [ \ell \mapsto \alpha ], \sigma'' \Downarrow_{pc}^E \sigma''', v}
          {(\code{let-label} ~ \ell ~ e_1 ~ e_2), \rho, \sigma \Downarrow_{pc}^E \sigma''', v}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{Obs}]
                    {e_1, \rho, \sigma \Downarrow_{pc}^E \sigma', \ell
                      \and e_2, \rho, \sigma' \Downarrow_{pc}^E \sigma'', v  \\\\
                      \langle (\lambda x.~e), \rho' \rangle = \sigma''(\ell) \\\\
                      e, \rho' [x \mapsto v], \sigma'' \Downarrow_{pc}^E \sigma''', v^{\pm} \\\\
                      e_3, \rho, \sigma''' \Downarrow_{pc}^E \sigma'''', v' \\\\
                      v'' = \fcol{\var{obs}(\ell, v', v^{\pm})}}
                    {(\code{obs} ~ e_1 ~ e_2 ~ e_3), \rho, \sigma \Downarrow_{pc}^E \sigma'''', v''}}
    \end{array}
    \]
    \\
    \hfill \textit{(Possibly-Faceted Application)} \boxed{\fcol{(v_1 ~ v_2), \rho, \sigma \Downarrow_{pc}^A \sigma, v}}
    \\
    \begingroup
    \color{\colorMATH}
    \begin{displaymath}
      \begin{array}{c}
        \begin{array}{ccc}
          {\color{\colorMATH}
          \inferrule[\textsc{App-$\star$}]
          { }
          {(\star ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma, \star}}
          & 
          {\color{\colorMATH}
          \inferrule[\textsc{App-Base}]
          {e, \rho'[x \mapsto v], \sigma \Downarrow_{pc}^E \sigma', v' }
          {(\langle \lambda x.~e, \rho' \rangle ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{App-Split}]
          {\{+\ell, -\ell\} \cap pc = \varnothing \and
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc \cup \{+\ell\}}^A \sigma', {v^+}' \\\\
            (v^- ~ v), \rho, \sigma' \Downarrow_{pc \cup \{-\ell\}}^A \sigma'', {v^-}' \and
            v' = \formfacet{\{+ \ell\}}{{v^+}'}{{v^-}'}}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma'', v'}}
        \end{array}
        \\
        \begin{array}{cc}
        \\\\
          {\color{\colorFAC}
          \inferrule[\textsc{App-Facet-Pos}]
          {+\ell \in pc \\\\
            (v^+ ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
          & 
          {\color{\colorFAC}
          \inferrule[\textsc{App-Facet-Neg}]
          {-\ell \in pc \\\\
            (v^- ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}
          {(\facet{\ell}{v^+}{v^-} ~ v), \rho, \sigma \Downarrow_{pc}^A \sigma', v'}}
        \end{array}
      \end{array}
    \end{displaymath}
    \endgroup
    \caption{Semantics of Faceted Execution}
    \label{fig:semantics}
\end{figure*}

Our semantics is shown in Figure~\ref{fig:semantics}. As faceted
execution is an extension of the lambda calculus with references, we
present the parts of the semantics unique to faceted execution in
red. Base values in our semantics include addresses (used for boxes),
constants, closures. We also include a special value \bmth{\star},
which represents lazy failure: \bmth{\star} helps define the semantics
of store update within a privileged context.

Values in our semantics are either (unfaceted) base values or facets
composed of a label and two branches. Facets can nest, allowing trees
of faceted values. We define \emph{branches} as positive or negative
labels. The program counter \rmth{pc} tracks the set of labels in the
execution context. For example, to apply a faceted function to a value
(for example \bmth{(\facet{\ell}{\lambda x.~0}{\lambda x.~1} 1)}), the
semantics first applies the function along the positive branch
extending the pc with \bmth{+\ell}, then applies the negative branch
extending the pc with \bmth{-\ell}.

The reduction relation \bmth{e, \rho, \sigma \Downarrow_{pc}^E \sigma,
  v} reduces an expression, environment, and store to a resulting
store and value. The first three rules (in all blue) are unchanged
from the standard interpretation in the lambda calculus. The
\textsc{Apply} rule calls out to the helper relation \rmth{(v~v),
  \rho, \sigma \Downarrow_{pc}^A \sigma, v}, which applies a
possibly-faceted value to an argument: if the value being applied is a
plain (unfaceted) closure, the \textsc{App-Base} rule (in blue, as it
is unchanged from the lambda calculus) applies it and returns
immediately to the \bmth{\Downarrow_{pc}^E} relation.

In the case that a faceted value is applied, \rmth{\Downarrow_{pc}^A}
performs one of three functions, based on the relation of \rmth{\ell}
to \rmth{pc}. If there is no occurrence of either \rmth{+\ell} or
\rmth{-\ell} in \rmth{pc}, then the semantics must split the
application. To do this, it applies both the positive and negative
branches after extending \rmth{pc}. After reducing both branches to a
value, the result is then formed into a facet. If \rmth{+\ell \in pc},
then the semantics has already branched on the label \rmth{\ell}, so
splitting would be redundant. In this case, \rmth{\Downarrow_{pc}^A}
simply selects the positive branch to apply and continues without
splitting. The symmetric case occurs in \textsc{App-Facet-Neg}. Facet
formation follows this pattern, accounting for the relation of
\rmth{\ell} to \rmth{pc}.

The rule \text{Box}, \textsc{Unbox}, and \textsc{Set} are similar to
the standard implementation of boxes, but with a slight twist: box
creation, reads, and writes may occur within a privilaged context. To
understand why, consider the following example\footnote{Our formal
  semantics elides \code{if}, though it may be obtained via a Church
  encoding if desired as in Austin et al.\cite{Austin:12}. Our
  implementation includes \code{if} but not other constructs such as
  \code{cond}}:

\begin{lstlisting}[language=Racket,escapechar=|,name=boxexample]
(define x (box 0))
(if (= (facet alice 0 1) 0)
  (set! x 0)
  (set! x 1))
(unbox x)
\end{lstlisting}

If we do nothing special to account for the fact that the program
branches on the facet, control flow implicitly launders the value
through the box to an unfaceted value. To fix this, we form a facet by
taking into account \rmth{pc} and forming a facet using the
meta-function \rmth{\formfacet{\ell}{v^+}{v^d}}. This meta-function is
defined in Figure~\ref{fig:metafunctions}, and takes three arguments:
the current \rmth{pc}, a positive view (\rmth{v^+}), and a ``default''
view (\rmth{v^d}). Facet construction builds a facet with a spine
corresponding to all of the labels in \rmth{pc}, and inserts
\rmth{v^+} at the focus corresponding to \rmth{pc}, putting the
default value \rmth{v^d} along all other branches. In the \code{box}
form, the facet uses a default value of \bmth{\star}. In terms of our
above example, this means that along the positive branch \code{x}
would be set to \facet{\code{alice}}{1}{\star} (as the false branch of
the \code{if} is taken), and along the subsequent negative branch
\code{x} is extended to \facet{\code{alice}}{1}{0}.

@@

Label creation allocates a label as a fresh address in the store,
binding the specified label predicate and adding it to the
environment. Observation evaluates the label expression to an address
and executes the associated predicate. Once this is done, \textsc{Obs}
uses the \rmth{\var{obs}} meta-function to select the appropriate branch
based on the value returned by the predicate associated with the
label.
